Revisiting `Growth and Inequality in Public Good Provision' ---Reproducing and Generalizing Through Inconvenient Online Experimentation

### 2023-06-16 analysis/article/paper.tex
rewrite transparency chapter

\begin{abstract}
I revisit the dynamic public goods game developed by Gächter et
al.~(2017) to study cooperation under dynamic interdependencies.
Collecting data from both a convenient (students) and an inconvenient
(general population) sample, I not only reproduce some of the authors'
original observations but also test their novel game's generalizability.
Appending a charitable dictator game, I find no correlations between
behavior in the charitable context and the dynamic game. This applies to
students and the general population sample alike. Because the study of
inexperienced general population samples raises methodological
challenges, such as fatigue and dropouts, this research approaches them.
Doing so, I provide simple solutions to run reliable interactive
experiments online.
\end{abstract}

@@

\begin{keyword}
    Public goods \sep Dynamic
games \sep Reproducibility \sep Generalizability \sep Online
Experiment \sep 
    Climate change mitigation
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

Today's actions are tomorrow's result. There are many settings in which
current decisions affect future outcomes and with it, future decision
spaces. Opting for environmental friendly policies today not only
reduces carbon dioxide omissions immediately but also helps us to reach
the Paris climate targets tomorrow. Deferring these policies, may not
necessarily prevent us from reaching these targets, but it requires more
effort in the future compared to a path that includes immediate action
\citet{HaenselEtAl2022}. Aiming at certain goals, today's actions (or
the omission thereof) not only affect intermediate outcomes but also the
number of paths one can choose from that lead to that specific goal.

@@

A game designed by Gächter, Mengel, Tsakas \& Vostroknutov \citeyearpar[
hereafter, GMTV]{GMTV2017} as well as Stefan Große (unpublished) shows
that it is fairly simple to add this element of realism. They
incorporate interdependencies into a \emph{dynamic} public goods game
(dPGG) by defining endowments as the income of previous periods.
Consequently, participants' actions in a given period affect their
number of actions in subsequent periods: the more (less) they earn now,
the more (less) they can contribute in the next period. Importantly,
this modification qualitatively yields the same rational predictions as
the static game (i.e.~free-riding and the under-provision of the public
good). It is thus, equally well suited to study dilemma situations.

Because there is surprisingly little experimental research on
interdependencies\footnote{One exemption are \citet{Moser2019}, who
  built on GMTVs' design to investigate leadership.}, I reproduced one
of GMTVs' treatments to compare dynamics across (in)experienced samples
to investigate its generalizability. \citet{GKLS2020} find that static
public goods games do not generalize well to real-world climate action.
They also find that generalizability depends on structural resemblance
of the public goods game with the context of climate change mitigation:
Greater resemblance improves generalizability. Because GMTVs' dynamic
setting has a more realistic property---namely, interdependencies---one
would expect it to be better suited to inform public policy. To test
this intuition, I not only ran the experiment with different samples but
also observed the participants' behavior in voluntary climate actions
(VCA).\footnote{The VCA is a (charitable) dictator game where each
  participant is a dictator dividing her budget between herself and some
  organization linked to the reduction of CO2 emissions.
  \citet{EckelGrossman1996} were the first to implement a charitable
  dictator game observing contributions of 30\% of the endowments. Like
  \citet{GKLS2020}, \citet{CarpenterEtAl2008} report that students make
  lower contributions to charity than community members.} This yielded a
setting similar to the one of \citet{GKLS2020}, which allows me to
analyze how behavior in the abstract game translates into real-world
action across samples. This research shows that the dynamic setting
\emph{does not} add any advances with respect to generalizability of
results.

\citet{AGM2018} conducted static public goods games in the lab and on
MTurk to draw (and report) lessons from online experimentation. This
study extends this literature \citep[see,
e.g.,][]{SnowbergYariv_2021, GuptaRigottiWilson_2021, GoodmanPaolacci2017, AmirEtAl2012}
by focusing on an \emph{inconvenient} sample (i.e.~one that is
completely inexperienced sample that has not been exposed to interactive
experiments before) playing a computationally more complex game. This
required me to design a robust (and thus, more complex) software to
minimize attrition. I collected paradata---which captures screen time
and survey navigation, for instance---to assess the desired fluency and
feasibility of the experiment. Like \citet{AGM2018}, this research is of
practical relevance as it reports on the robust design that made the
dynamic game feasible.

@@

After commenting on transparent research practices and reporting the
methods in Section~\ref{sec-transparency} and Section~\ref{sec-methods},
this paper is organized along these findings: The confirmatory
reproduction is reported in Section~\ref{sec-replication}, the online
feasibility in Section~\ref{sec-feasibility} and the generalizability in
Section~\ref{sec-generalizability}. Section~\ref{sec-conclusion}
concludes. Because I observed some (qualitatively inconsequential)
differences in reproducing GMTVs' analysis using their data, I report
these in the \href{@sec-appendix}{Appendix}.

@@

Well before credibility crises comprised a variety disciplines,
\citet[p.~2]{bemwriting} indirectly conceded discrepancies between fuzzy
research processes and the polished article that results from it:

@@

Now, 20 years later, we accustom ourselves with mechanisms designed to
unravel the fuzzy back-and-forth between exploratory and confirmatory
research. Prominently, pre-registrations were established to tie the
researchers' hands when doing confirmatory research--reducing observable
discrepancies between Daryl Bem's above-mentioned (a) and (b). Today,
researchers can not only choose where to pre-register but also decide
about the level of detail they provide. More severely, nothing stops
researchers from pre-registering multiple hypotheses opposing each other
\citetext{\citealp[see, e.g.,][]{SimmonsEtAl2021}; \citealp[and][ for a
discussion]{PhamEtAl2021}}.

\citet{WaldronAllen2022} report a wide variety of depth of
pre-registrations in the field of Neuropsychopharmacology and document
that the \emph{`pre-registered'} label can be obtained with minimal
effort. In absence of transparency and scrutiny pre-registrations run
into danger to miss the target of making science more credible. Even
more so, they may have adverse effects as the \emph{`pre-registered'}
label becomes cheap (too) to acquire.

Discussing the costs and benefits of pre-registrations,
\citet{Olken2015} stresses the uncertainty of research processes and
promotes moderate approaches. As \citet{WaldronAllen2022}, he argues
that pre-registrations not necessarily eliminate flexibility. These
`moderate' or `flexible' approaches require transparency to reduce
concerns of \emph{`data-mining'} \citep[ p.~61]{Olken2015} by showing
(a) the analyses I planned to run, (b) the analysis that made most sense
after collecting the data and the transition from (a) to (b). Because
this is neither incentivized nor easy to verify, we need tools that make
transparency convenient to establish and to scrutinize. I propose
version control (such as Git) as one such tool and illustrate its
application on this paper: GitHub is a website and cloud-based service
where developers---and researchers alike---can store and manage their
code. The service is based on Git\footnote{Git is a specific open-source
  version control system developed in 2005.} and designed for version
control. Importantly, changes are time-stamped and can be tracked.
Moreover, one can create branches, that is, duplicates of code either to
work collaboratively or to archive a certain state.

To sum up, pre-registrations alone are necessary but not sufficient to
restore the credibility in experimental sciences. Pre-registrations must
come with additional emphasis on transparency covering the whole
research process as well as with a higher level of scrutiny that looks
beyond mere labels.

In fact, the experiment reported in this article qualifies for the
\emph{`pre-registered'} label at a first glance: The analysis was
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. Further, I pre-registered the exact analyses I
planned to run when I designed the experiment on
\href{https://shorturl.at/bAGKY}{GitHub}.\footnote{https://shorturl.at/bAGKY
  to run the code, you need to execute the .Rmd files in this repository
  in the order that is indicated by file names.}

At a second glance, the analysis that made the most sense after having
collected the data and that is reported here is a little different
though: Because the more representative subject pool was exhausted
earlier than expected, I recruited students which opened up a new
research direction: assessing generalizability. Hence, this article
combines both confirmatory as well as exploratory research.

Accordingly, I edited many of the analyses scripts.\footnote{The source
  code of this document contains the analysis and can be found
  \href{https://github.com/Howquez/coopUncertainty/blob/main/analysis/quarto/paper.qmd}{here}:
  https://shorturl.at/ilpAG} However, because the originally planned
analysis code is archived and reproducible, these changes are
transparent. Importantly, I followed a literate programming approach
\citep{Knuth_1984, AkhtarYe_2023}. Hence, all documents needed to
analyze the data and to write the report stem from the same source. This
establishes consistence between the commands one tells the statistical
software to do and the explanation one tells human beings one told the
statistical software to do. As such, (a), (b) and the transition from
(a) to (b) are not only transparent but also comprehensible.

@@

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific} reproduction\footnote{\citet{Parsons2022}
  use the term of a \emph{conceptual replication} which means the same.}
of one treatment of GMTVs' dynamic public goods game. The pure
reproduction re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
reproduction, where I utilize a different sample drawn from a different
population in a different situation, is described in the following
sections.

@@

Considering the design, the only differences between such a
\emph{static} game (with partner matching) and GMTVs' design are
\emph{dynamics}: Instead of receiving fresh endowments every period,
participants receive one endowment only at the beginning of the first
period. A participant's endowment in the second period is the wealth she
accumulated in the first period. A participant's endowment in the third
period is the wealth she accumulated in the first two periods. And so
on. Hence, a decision in one period has consequences on future
endowments and, ultimately, growth paths. For this reason, the game is
described as a \emph{dynamic} public goods game.

@@

Like GMTV I employ a real giving task after the abstract game.
Specifically, I employed a (charitable) dictator game where each
participant is a dictator dividing her budget between herself and some
organization. Whereas GMTV chose \emph{Doctors without Borders} as a
receiver, I chose another organization in another context: like
\citet{GKLS2020}, I employed a VCA, where participants could donate any
amount of their earnings to offset carbon dioxide (that is, retire
emission permits from the EU ETS).\footnote{Importantly,
  \citet{GKLS2020} made the VCA decision with a fresh endowment
  \emph{before} they played the abstract game. I deviate from their
  procedure to match GMTVs' procedure.} To ensure that each participant
had the same basic level of information about the impact of their
decision, I provided some basic information about the mechanism. The
information also highlighted that the mitigation came into operation on
a European level. Finally, I informed the participants that the
documentation of individual and aggregate contributions were to be
posted immediately after the concluding the sessions online. To avoid
privacy or social image concerns, participants learned their unique and
random IDs, which they needed to identify their individual
contributions. The document certified that their contributions have been
used to offset
\href{https://www.compensators.org/compensatelist/?searchterm=stefan+traub}{1.82
tons} of carbon dioxide emissions.

@@

I recruited the participants from the so called
\emph{\href{https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html}{HamburgPanel}}
using HROOT \citep{hroot}. The panel is provided by the University of
Hamburg's Research Laboratory, which used a randomized last digits
approach to build the panel while drawing from the population of
citizens of Hamburg, Germany. Because the sample was exhausted at one
point, I also recruited students from the University of Hamburg.

@@

Throughout this paper, I will compare results of my experiment with the
results of GMTVs' NOPUNISH 10 Period treatments. I am thus, referring to
three different samples utilized at two points in time: the University
of Nottingham's students (in late 2012), Hamburg's citizens and the
University Hamburg's students (both in July 2021).
Table~\ref{tbl-sample-properties} gives an overview.

@@

The experiment was logistically complex for several reasons. First, the
sample was inexperienced. Second, the experiment was interactive and
synchronous. Third, the underlying game was dynamic and interdependent.
This makes dropouts not only more likely but also more expensive, which
is why attrition was a major concern implementing the experiment.

I chose oTree \citep{oTree} to implement the experiment because it is
open-source, well documented and very flexible. Its
\href{https://getbootstrap.com/}{Bootstrap} (a powerful frontend
toolkit) integration allowed me to make the graphical user interface
interactive, appealing and easy to navigate. The
\href{https://www.highcharts.com/}{Highcharts library} made it easy to
visualize results and to communicate dynamics. Insofar, oTree served a
good tool to enhance the participants' user experience and thus, to make
dropouts less likely.

Which features were required to handle dropouts? First, participants had
to be matched to form a group \emph{after} comprehension questions were
answered successfully. Importantly, participants were grouped by the
order they answered these questions to reduce waiting times. While
waiting for other players to form the group, the participants saw a
wait-page informing them that they are waiting for other participants to
arrive and that they do not have to wait for longer than 10 minutes. The
screen also informed them that they would receive a \emph{patience
bonus} of one Euro after the expiration of that time (or what was left
of it). Second, participants only had 10 minutes to make the first
contribution and 4 minutes for the remaining contributions. After this
time expired, participants were replaced by bots that made random
contributions. In this case, the remaining group members were informed
about the replacement. Both features were implemented to limit wait
times and boredom for other participants. Section~\ref{sec-feasibility}
shows that the first feature became effective in some cases, wheres the
second feature did not.

@@

Participants entered the experiment at appointed times remotely from
home. They first saw a welcome screen. After agreeing to the privacy
policy, they could proceed to the instructions individually. Having read
these instructions, each participant has also seen a demo-screen
explaining the user interface. Before proceeding, they had to answer six
comprehension questions correctly to avoid confusion in later stages
\citep{FerraroVossler2010}. Subsequently, they saw a waiting screen
until they could be matched with three other participants, who have
answered the comprehension questions correctly. Once matched, they were
exposed to the decision screen over ten periods. At the end of the last
period, participants saw results of all periods. Subsequently, they made
their VCA decision, before I elicited risk preferences
\citep{HoltLaury2002} and finished with GMTVs' questionnaire.

@@

The experiment lasted around 25 minutes on average. The earnings
averaged 11.23 Euros (sd = 4.85).\footnote{This values include earnings
  from the incentivized risk elicitation task that is not part of the
  analysis.}

@@

\hypertarget{sec-replication}{%
\subsection{Pre-registered GMTV Reproduction}\label{sec-replication}}

@@

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is the reproduction sample (consisting
of both students and non-students) more pro-social than the original
sample? A two-sided rank sum test reveals that it is not (p=0.393). Both
samples contributed 10 tokens, that is, 50\% of their endowments on
average (median and mean). Moreover, both samples' initial contributions
resemble initial contributions participants usually make in the static
game with partner matching.\footnote{See Figure 3B in
  \citet[p.989]{fehrgaechter2000}, for instance.} However, in the
dynamic game presented here, I are particularly interested in the
subsequent periods because differences add up exponentially. Do the two
groups remain similar over the course of time?

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}.
Figure~\ref{fig-share-of-contributions} illustrates that the samples
make similar contributions at the beginning and the end of the game but
behave differently in between. More precisely, the left panel--depicting
the average contributions in absolute terms--shows that the original
sample contributed more than the reproduction sample \emph{in all but
the first and last period}.\footnote{In period five, this difference is
  significant: a two-sided rank sum test yields p=0.0486.} For this
reason, the original sample's behavior differs from the reproduction
sample's behavior in two aspects: it contributes more and exhibits a
considerable drop in the last period (whereas the reproduction sample's
contributions flatten).

@@

Again, both samples' behavior resembles the contributions participants
usually make in the static game with partner matching: contributions
equal approximately half of endowments in the very first period and
decrease to around ten percent of endowments by the last
period.\footnote{The right panel is thus, comparable to the
  visualizations \emph{and results} in the static game. See, for
  instance, Figure 1B in \citet[p.986]{fehrgaechter2000}.} In the
dynamic game presented here, however, different paths lead to different
levels of wealth -- even if they share the same start- and end-points. I
am thus, more interested in the contributions' implications for wealth
generation and growth.

@@

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more, one would expect the respective groups
to be more wealthy. A mere mean comparison indicates just that: An
average group in the original sample accumulated about 478 tokens. In
contrast, an average group in the reproduction sample accumulated about
380 tokens. This difference is insignificant at conventional levels
though: A two-sided rank sum test (comparing differences between
samples) yields p=0.136 for the mean stock in last period of the game.

Although there clearly is growth, groups do not realize the maximal
potential efficiency: under full cooperation, a group can accumulate at
maximum 4613 tokens or EUR 230. This is depicted in the left panel of
Figure~\ref{fig-growth-heterogeneity}, where one can see the average
wealth over time by sample. The panel illustrates for both samples that
growth was continuous and surprisingly linear, given the exponential
character of the game's design. Despite somewhat differing contribution
behavior between samples, neither the eventual wealth nor the
corresponding growth paths differed. Differences in contribution
behavior did, thus, not translate to significantly different wealth
outcomes.

@@

\textbf{Result 1.} \emph{The \texttt{NOPUNISH\ 10} treatment of GMTV can
be replicated because the replication data resemble the original data
with respect to initial and final contributions, wealth and growth as
well as inequality.}

>>>>

\begin{abstract}
I revisit the dynamic public goods game developed by Gächter et
al.~(2017) to study cooperation under dynamic interdependencies.
Collecting data from both a convenient (students) and an inconvenient
(general population) sample, I not only reproduce some of the authors'
original observations but also test their novel game's generalizability.
Appending a charitable dictator game, I find no correlations between
behavior in the charitable context and the dynamic game. This applies to
students and the general population sample alike. Because the study of
inexperienced general population samples raises methodological
challenges, such as fatigue and dropouts, this research approaches them.
In doing so, I provide simple solutions to run reliable interactive
experiments online. Furthermore, this article showcases the use of
literate programming and version control which I argue are convenient
tools to make pre-registrations more credible and flexible.
\end{abstract}

@@

\begin{keyword}
    Public goods \sep Dynamic
games \sep Reproducibility \sep Generalizability \sep Online
Experiment \sep 
    Climate change mitigation
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, sharp corners, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

Today's actions are tomorrow's result. There are many settings in which
current decisions affect future outcomes and with it, future decision
spaces. Opting for environmentally friendly policies today not only
reduces carbon dioxide emissions immediately but also helps us to reach
the Paris climate targets tomorrow. Deferring these policies, may not
necessarily prevent us from reaching these targets, but it requires more
effort in the future compared to a path that includes immediate action
\citep{HaenselEtAl2022}. Aiming at certain goals, today's actions (or
the omission thereof) not only affect intermediate outcomes but also the
number of paths one can choose from that lead to that specific goal.

@@

A game designed by Gächter, Mengel, Tsakas \& Vostroknutov \citeyearpar[
hereafter, GMTV]{GMTV2017} as well as Stefan Große (unpublished), shows
that it is fairly simple to add this element of realism. They
incorporate interdependencies into a \emph{dynamic} public goods game
(dPGG) by defining endowments as the income of previous periods.
Consequently, participants' actions in a given period affect their
number of actions in subsequent periods: the more (less) they earn now,
the more (less) they can contribute in the next period. Importantly,
this modification qualitatively yields the same rational predictions as
the static game (i.e.~free-riding and the under-provision of the public
good). It is thus, equally well suited to study dilemma situations.

Because there is surprisingly little experimental research on
interdependencies\footnote{See e.g. \citet{BattagliniEtAl2016},
  \citet{Rockenbach2017} or \citet{Moser2019}.}, I reproduced one of
GMTVs' treatments to compare dynamics across (in)experienced samples to
investigate its generalizability. \citet{GKLS2020} find that static
public goods games do not generalize well to real-world climate action.
They also find that generalizability depends on the structural
resemblance of the public goods game with the context of climate change
mitigation: Greater resemblance improves generalizability. Because
GMTVs' dynamic setting has a more realistic property---namely,
interdependencies---one would expect it to be better suited to inform
public policy. To test this intuition, I not only ran the experiment
with different samples but also observed the participants' behavior in
voluntary climate actions (VCA).\footnote{The VCA is a (charitable)
  dictator game where each participant is a dictator dividing her budget
  between herself and some organization linked to the reduction of CO2
  emissions. \citet{EckelGrossman1996} were the first to implement a
  charitable dictator game observing contributions of 30\% of the
  endowments. Like \citet{GKLS2020}, \citet{CarpenterEtAl2008} report
  that students make lower contributions to charity than community
  members.} This yielded a setting similar to the one of
\citet{GKLS2020}, which allows me to analyze how behavior in the
abstract game translates into real-world action across samples. This
research shows that the dynamic setting \emph{does not} add any advances
to the generalizability of results.

\citet{AGM2018} conducted static public goods games in the lab and on
MTurk to draw (and report) lessons from online experimentation. This
study extends this literature \citep[see,
e.g.,][]{SnowbergYariv_2021, GuptaRigottiWilson_2021, GoodmanPaolacci2017, AmirEtAl2012}
by focusing on an \emph{inconvenient} sample (i.e.~one that is a
completely inexperienced sample that has not been exposed to interactive
experiments before) playing a computationally more complex game. This
required me to design robust (and thus, more complex) software to
minimize attrition. I collected para data---which capture screen time
and survey navigation, for instance---to assess the desired fluency and
feasibility of the experiment. Like \citet{AGM2018}, this research is of
practical relevance as it reports on the robust design that made the
dynamic game feasible.

@@

After commenting on transparent research practices and reporting the
methods in Section~\ref{sec-transparency} and Section~\ref{sec-methods},
this paper is organized along these findings: The confirmatory
reproduction is reported in Section~\ref{sec-replication}, the online
feasibility in Section~\ref{sec-feasibility}, and the generalizability
in Section~\ref{sec-generalizability}. Section~\ref{sec-conclusion}
concludes. Because I observed some (qualitatively inconsequential)
differences in reproducing GMTVs' analysis using their data, I report
these in the \href{@sec-appendix}{Appendix}.

@@

Well before credibility crises comprised a variety of
disciplines\footnote{See e.g., \citet{CamererEtAl2016};
  \citet{BrodeurEtAl2016}; \citet{BrodeurEtAl2020};
  \citet{ChristensenMiguel2018}; \citet{FerraroShukla2020} who focus on
  economics.}, \citet[p.~2]{bemwriting} indirectly conceded
discrepancies between fuzzy research processes and the polished article
that results from it:

@@

Now, 20 years later, we accustom ourselves to mechanisms designed to
unravel the fuzzy back-and-forth between exploratory and confirmatory
research. Prominently, pre-registrations were established to tie the
researchers' hands with respect to p-hacking and ex-post theorizing when
doing confirmatory research---reducing observable discrepancies between
Daryl Bem's above-mentioned (a) and (b).

While pre-registrations could help to make empirical sciences more
credible, they come at costs that are discussed by \citet{Olken2015} as
well as \citet{Krishna2021}, \citet{SimmonsEtAl2021}, and
\citet{PhamEtAl2021}. One important concern is that strict
pre-registrations may severely discount any additional analysis of the
data inspired by surprise results. In practice, there are at least two
ways to approach that problem: First, one could (and should) add
unforeseen analyses to the article but label them as exploratory.
Second, one could write a minimal pre-registration leaving enough
degrees of freedom for the researcher to incorporate such surprises
without further notice. The key difference between the two approaches is
the deliberate omission of transparency.

For the field of neuropsychopharmacology \citet{WaldronAllen2022}
document that the depth of pre-registrations differs indeed: it is
spanning from notes on hypotheses only to fully specified, detailed
experimental protocols with accompanying power analyses and sampling
plans. Treating all pre-registrations equally may thus lead to an
adverse effect: If one can acquire the \emph{`pre-registered'} label
with minimal effort and maximum flexibility, the label itself may run
into danger to become worthless and to miss the target of making science
more credible.

This erosion of credibility can be avoided by the means of
transparency---not because transparency can tighten the ties but also
because it facilitates scrutiny, policing
\citep[see][]{AnkelPetersFialaNeubauer2023} and forensics: transparent
research practices make it easier for others to track and understand a
researcher's reaction to surprises in the data, which, in turn, gives
the researcher more flexibility to react to these surprises. As
additional transparency is associated with additional effort, I propose
version control as a natural and easy-to-implement solution augmenting
pre-registrations.

Contemporary experimental research is, to a certain degree, a
computational science: experiments often require complex software and
the analysis thereof can result in complicated scripts. In addition,
much research is carried out in collaborative efforts. For these reasons
alone, it makes sense to implement version control systems such as
GitHub and GitLab as well as OSF (albeit less powerful) that ensure that
files are stored consistently and that changes are tracked. When these
systems are implemented in the whole research process, one can archive
different states of the software and visualize differences between
different states. For instance, one can archive the analysis before and
after the data was collected or the experimental software used in a
pilot and the eventual experiment. Importantly, one can keep projects
private and change their visibility to public only after the
corresponding article is accepted for publication. Taken together, these
features \emph{automatically} enhance transparency and facilitate
scrutiny, policing, and forensics whilst also making software
development and collaboration more efficient. As a consequence, the
implementation and eventual disclosure of version control come not only
with meta-scientific but also with organizational benefits.

The experiment reported in this article qualifies for the
\emph{`pre-registered'} label at first glance: The analysis was
pre-registered in the American Economic Association's RCT Registry
\citep{preregistration}. Further, I pre-registered the exact analyses I
planned to run when I designed the experiment on
\href{https://github.com/Howquez/GMTV/tree/July21Replication/analysis/R}{GitHub}.\footnote{bit.ly/3Pnak5H.
  To run the code, you need to execute the .Rmd files in this repository
  in the order that is indicated by file names. Detailed instructions
  can be found in the README file accompanying the analysis scripts.}

At second glance, the analysis that made the most sense after having
collected the data and that is reported here is a little different
though: Because the more representative subject pool was exhausted
earlier than expected, I recruited students which opened up a new
research direction: assessing generalizability. Hence, this article
combines both confirmatory as well as exploratory research.

Accordingly, I edited many of the analysis scripts.\footnote{The source
  code of this document contains the analysis and can be found
  \href{https://github.com/Howquez/GMTV/blob/main/analysis/article/paper.qmd}{here}:
  https://bit.ly/43LCyLw} However, because the originally planned
analysis code is archived and reproducible, these changes are
transparent. Importantly, I followed a literate programming approach
\citep{Knuth_1984, AkhtarYe_2023}. Hence, all documents needed to
analyze the data and to write the report stem from the same source. This
establishes consistence between the commands one tells the statistical
software to do and the explanation one tells human beings one told the
statistical software to do. As such, (a), (b), and the transition from
(a) to (b) are not only transparent but also comprehensible.

@@

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific} reproduction\footnote{In fact,
  \citet{Hamermesh2007} uses the term \emph{replication} instead of
  \emph{reproduction}. The reason I deviate is that I do not compare
  different treatments with each other to \emph{replicate} a treatment
  effect. Instead, I \emph{reproduce} the results of one specific
  treatment arm.} of one treatment of GMTVs' dynamic public goods game.
The pure reproduction re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{the errors I identified} in the original paper. The scientific
reproduction, where I utilize a different sample drawn from a different
population in a different situation, is described in the following
sections.

@@

Considering the design, the `only' differences between such a
\emph{static} game (with partner matching) and GMTVs' design are
\emph{dynamics}: Instead of receiving fresh endowments every period,
participants receive one endowment only at the beginning of the first
period. A participant's endowment in the second period is the wealth she
accumulated in the first period. A participant's endowment in the third
period is the wealth she accumulated in the first two periods. And so
on. Hence, a decision in one period has consequences on future
endowments and, ultimately, growth paths. For this reason, the game is
described as a \emph{dynamic} public goods game.

@@

Like GMTV I employ a real giving task after the abstract game.
Specifically, I employed a (charitable) dictator game where each
participant is a dictator dividing her budget between herself and some
organization. Whereas GMTV chose \emph{Doctors without Borders} as a
receiver, I chose another organization in another context: like
\citet{GKLS2020}, I employed a VCA, where participants could donate any
amount of their earnings to offset carbon dioxide (that is, retire
emission permits from the EU ETS).\footnote{Importantly,
  \citet{GKLS2020} made the VCA decision with a fresh endowment
  \emph{before} they played the abstract game. I deviate from their
  procedure to match GMTVs' procedure.} To ensure that each participant
had the same basic level of information about the impact of their
decision, I provided some basic information about the mechanism. The
information also highlighted that the mitigation came into operation on
a European level. Finally, I informed the participants that the
documentation of individual and aggregate contributions was to be posted
immediately after the conclusion of the sessions online. To avoid
privacy or social image concerns, participants learned their unique and
random IDs, which they needed to identify their individual
contributions. The document certified that their contributions have been
used to offset
\href{https://www.compensators.org/compensatelist/?searchterm=stefan+traub}{1.82
tons} of carbon dioxide emissions.

@@

I recruited the participants from the so-called
\emph{\href{https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html}{HamburgPanel}}
using HROOT \citep{hroot}. The panel is provided by the University of
Hamburg's Research Laboratory, which used a randomized last digits
approach to build the panel while drawing from the population of
citizens of Hamburg, Germany. Because the sample was exhausted at one
point, I also recruited students from the University of Hamburg.

@@

Throughout this paper, I will compare the results of my experiment with
the results of GMTVs' NOPUNISH 10 Period treatments. I am thus,
referring to three different samples utilized at two points in time: the
University of Nottingham's students (in late 2012), Hamburg's citizens,
and the University of Hamburg's students (both in July 2021).
Table~\ref{tbl-sample-properties} gives an overview.

@@

The experiment was logistically complex for several reasons. First, the
sample was inexperienced. Second, the experiment was interactive and
synchronous. Third, the underlying game was dynamic and interdependent.
This makes dropouts not only more likely but also more expensive, which
is why attrition was a major concern in implementing the experiment.

I chose oTree \citep{oTree} to implement the experiment because it is
open-source, well-documented, and very flexible. Its
\href{https://getbootstrap.com/}{Bootstrap} (a powerful frontend
toolkit) integration allowed me to make the graphical user interface
interactive, appealing, and easy to navigate. The
\href{https://www.highcharts.com/}{Highcharts library} made it easy to
visualize results and communicate dynamics. Insofar, oTree served as a
good tool to enhance the participants' user experience and thus, to make
dropouts less likely.

Which features were required to handle dropouts? First, participants had
to be matched to form a group \emph{after} comprehension questions were
answered successfully. Importantly, participants were grouped by the
order they answered these questions to reduce waiting times. While
waiting for other players to form the group, the participants saw a
wait-page informing them that they are waiting for other participants to
arrive and that they do not have to wait for longer than 10 minutes. The
screen also informed them that they would receive a \emph{patience
bonus} of one Euro after the expiration of that time (or what was left
of it). Second, participants only had 10 minutes to make the first
contribution and 4 minutes for the remaining contributions. After this
time expired, participants were replaced by bots that made random
contributions. In this case, the remaining group members were informed
about the replacement. Both features were implemented to limit wait
times and boredom for other participants. Section~\ref{sec-feasibility}
shows that the first feature became effective in some cases, whereas the
second feature did not.

@@

Participants entered the experiment at appointed times remotely from
home. They first saw a welcome screen. After agreeing to the privacy
policy, they could proceed to the instructions individually. Having read
these instructions, each participant has also seen a demo-screen
explaining the user interface. Before proceeding, they had to answer six
comprehension questions correctly to avoid confusion in later stages
\citep{FerraroVossler2010}. Subsequently, they saw a waiting screen
until they could be matched with three other participants, who have
answered the comprehension questions correctly. Once matched, they were
exposed to the decision screen over ten periods. At the end of the last
period, participants saw the results of all periods. Subsequently, they
made their VCA decision, before I elicited risk preferences
\citep{HoltLaury2002} and finished with GMTVs' questionnaire.

@@

The experiment lasted around 25 minutes on average. The earnings
averaged 11.23 Euros (sd = 4.85).\footnote{This value include earnings
  from the incentivized risk elicitation task that is not part of the
  analysis.}

@@

\hypertarget{sec-replication}{%
\subsection{Reproducibility}\label{sec-replication}}

@@

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is the reproduction sample (consisting
of both students and non-students) more pro-social than the original
sample? A two-sided rank sum test reveals that it is not (p=0.393). Both
samples contributed 10 tokens, that is, 50\% of their endowments on
average (median and mean). Moreover, both samples' initial contributions
resemble the initial contributions participants usually make in the
static game with partner matching.\footnote{See Figure 3B in
  \citet[p.989]{fehrgaechter2000}, for instance.} However, in the
dynamic game presented here, I am particularly interested in the
subsequent periods because differences add up exponentially. Do the two
groups remain similar over the course of time?

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{yes and no}.
Figure~\ref{fig-share-of-contributions} and
Table~\ref{tbl-contribution-periods} illustrate that the samples make
similar contributions at the beginning and the end of the game but
behave differently in between. More precisely, the left panel--depicting
the average contributions in absolute terms--shows that the original
sample contributed more than the reproduction sample \emph{in all but
the first and last period}.\footnote{In period five, this difference is
  significant: a two-sided rank sum test yields p=0.0486.} For this
reason, the original sample's behavior differs from the reproduction
sample's behavior in two aspects: it contributes more and exhibits a
considerable drop in the last period (whereas the reproduction sample's
contributions flatten).

@@

Again, both samples' behavior resembles the contributions participants
usually make in the static game with partner matching: contributions
equal approximately half of the endowments in the very first period and
decrease to around ten percent of endowments by the last
period.\footnote{The right panel is thus, comparable to the
  visualizations \emph{and results} in the static game. See, for
  instance, Figure 1B in \citet[p.986]{fehrgaechter2000}.} In the
dynamic game presented here, however, different paths lead to different
levels of wealth -- even if they share the same start- and endpoints. I
am thus, more interested in the contributions' implications for wealth
generation and growth.

@@

How do the different contribution paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied, and redistributed).} Given that the
original sample contributed more, one would expect the respective groups
to be more wealthy. A mere mean comparison indicates just that: An
average group in the original sample accumulated about 478 tokens. In
contrast, the average group in the reproduction sample accumulated about
380 tokens. This difference is insignificant at conventional levels
though: A two-sided rank sum test (comparing differences between
samples) yields p=0.136 for the mean stock in the last period of the
game.

Although there clearly is growth, groups do not realize the maximal
potential efficiency: under full cooperation, a group can accumulate a
maximum of 4613 tokens or EUR 230. This is depicted in the left panel of
Figure~\ref{fig-growth-heterogeneity}, where one can see the average
wealth over time by sample. The panel illustrates for both samples that
growth was continuous and surprisingly linear, given the exponential
character of the game's design. Despite somewhat differing contribution
behavior between samples, neither the eventual wealth nor the
corresponding growth paths differed. Differences in contribution
behavior did, thus, not translate to significantly different wealth
outcomes.

@@

\textbf{Result 1.} \emph{The \texttt{NOPUNISH\ 10} treatment of GMTV can
be replicated because the replication data resemble the original data
with respect to initial and final contributions, wealth, and growth as
well as inequality.}
